---
title: "Practical Machine Learning Project"
author: "SM"
date: "Sunday, August 23, 2015"
output: html_document
---
# Executive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. Goal of this projuect is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Data Processing

The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

```{r}
#Commented as the below doesn't work when using Knit HTML
#download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
#download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
pml_train <- read.csv('pml-training.csv')
pml_test <- read.csv('pml-testing.csv')
dim(pml_train)
table(pml_train$classe)
```
# Exploratory Data Analysis
Partitioning the dataset, Creating training, test and validation sets
```{r}
library(caret)
library(ggplot2)
library(lattice)
library(kernlab)
library(randomForest)

trainidx <- createDataPartition(pml_train$classe,p=.9,list=FALSE)
traindata = pml_train[trainidx,]
testdata = pml_train[-trainidx,]
set.seed(32768)
nzv <- nearZeroVar(traindata)
trainnzv <- traindata[-nzv]
testnzv <- testdata[-nzv]
pml_testnzv <- pml_test[-nzv]

dim(trainnzv)
dim(testnzv)
dim(pml_testnzv)
train_fidx <- which(lapply(trainnzv,class) %in% c('numeric'))
trainnzv1 <- preProcess(trainnzv[,train_fidx], method=c('knnImpute'))
train_fidx
trainnzv1
pred1 <- predict(trainnzv1, trainnzv[,train_fidx])
pred_train <- cbind(trainnzv$classe,pred1)
names(pred_train)[1] <- 'classe'
pred_train[is.na(pred_train)] <- 0

pred2 <- predict(trainnzv1, testnzv[,train_fidx])
pred_test <- cbind(testnzv$classe, pred2)
names(pred_test)[1] <- 'classe'
pred_test[is.na(pred_test)] <- 0
pred_pmltest <- predict(trainnzv1,pml_testnzv[,train_fidx] )
dim(pred_train)
dim(pred_test)
dim(pred_pmltest)

```
# Data Modeling
We will use random forest as our model as implemented in the randomForest package 
```{r}
ftmodel <- randomForest(classe~.,data=pred_train)

pred_train1 <- predict(ftmodel, pred_train) 
print(table(pred_train1, pred_train$classe))
training <- as.data.frame(table(pred_train1, pred_train$classe))
pred_test1 <- predict(ftmodel, pred_test) 
print(table(pred_test1, pred_test$classe))
str(pred_pmltest)
pred_answers <- predict(ftmodel, pred_pmltest) 
pred_answers
```
# Results
```{r}
answers <- predict(ftmodel, pred_pmltest) 
answers

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

```


```
